Solutions to Integer Programming from Quantum Annealing

### 2020-06-01 paper/main.tex
[Debug] Syntax
push for runs

\documentclass[prd,twocolumn,tightenlines,preprintnumbers,showpacs,superscriptaddress,notitlepage,nofootinbib,eqsecnum,floatfix,longbibliography]{revtex4}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{bm,bbm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{dsfont}

@@

% disable subsections and subsubsections in the TOC
\makeatletter
\def\l@subsubsection#1#2{}
\makeatother

@@

\newcommand{\alert}[1]{\textbf{\color{red}{#1}}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}

\begin{abstract}
While quantum computing proposes promising solutions to computational problems not accessible with classical approaches, due to current hardware constraints, most quantum algorithms are not yet capable of computing systems of practical relevance, and classical counterparts outperform them.
To practically benefit from quantum architecture, one has to identify problems and algorithms with favorable scaling and improve on corresponding limitations depending on available hardware.
For this reason, we developed an algorithm that solves integer linear programming problems (ILP), a classically NP-hard problem, on a quantum annealer (QA), and investigated problem and hardware-specific limitations.
This work presents the formalism of how to map ILP problems to the annealing architectures.
It illustrates the effects of decorrelation and many-body localization, for the minimum dominating set problem, and compares annealing results against numerical simulations of the quantum architecture.
We find that the QA ILP algorithm outperforms random guessing but is limited to small problems and that annealing schedules can be adjusted to reduce the effects of decorrelation--which is qualitatively confirmed by simulations.
%Conclusion: What do these findings mean to us?
[Rephrase] We need better hardware to challenge heuristic solvers but we identified how to optimize hardware.
%Perspectives: What should be done next?
[Rephrase] Better understand correlations and localizations and improve simulation to scale to larger problems.

\end{abstract}

@@

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents

@@

%========================================================================================
\section{INTRODUCTION}
\label{sec:introduction}
%========================================================================================

Integer linear programming (ILP) is an integer optimization problem subject to inequality constraints and is a commonly tackled problem applicable to situations such as x~\cite{}, y~\cite{}, and the Minimum Dominating Set Problem (MDS)~\cite{}.
In general, however, ILP is classically NP-complete, and as a result, heuristic methods are employed~\cite{}.
Standard classical heuristic algorithms employ a greedy scheme which iteratively approximates optimal solutions--starting from a random initial guess, these algorithms apply locally optimal choices at each step.
The NP-hardness of ILP can be understood by realizing that while the solution to an $n$-dimensional linear problem can be obtained in polynomial time, the optimal integer solution in general must be found in the $2^n$ integer solutions which surround the real number solution.
While greedy algorithms do not guarantee optimal global solutions, they find approximate solutions in polynomial time, which can be utilized in further computations.
Due to the importance and difficulty, in this work, we present a method to obtain the optimal solution to ILP by employing methods of quantum annealing.

Quantum annealing solves a general quadratic binary optimization problem (QUBO) by slowly varying a time-dependent Hamiltonian~\cite{}.
Through the adiabatic theorem of quantum mechanics, the annealer is initially prepared in a trivial ground state while the final Hamiltonian encodes the solution to the ILP.
Due to the explosion in research efforts towards hardware implementations of quantum annealers and future improvements to the annealing schedule~\cite{}, mapping ILP to QUBO provides a path forward towards obtaining optimal solutions to the class of integer optimization problems~\cite{2018Glover}.

\alert{[Formulate: Decoherence \& Localization]}
Hardware suffers from...
For this reason one is not guaranteed to find optimal solutions

\alert{[Formulate: Simulations]}
To understand findings and limitations, we mimic the hardware

In section \ref{sec:methods} we propose an algorithm that maps ILP problems to QA hardware (\ref{sec:methods:ilp}) and present an example implementation in case of the Minimal Dominating Set problem (\ref{sec:methods:mds}).
We review the annealing process on the utilized hardware (\ref{sec:methods:annealing}) and describe the computational framework which we used to simulate the annealing process (\ref{sec:methods:simulation}).
Section \ref{sec:results} summarizes our findings and presents...
We conclude this report in \ref{sec:conclusion}.
All algorithms used and data obtained are publicly available (\ref{sec:open-source})

%========================================================================================
\section{METHODS}
\label{sec:methods}
%========================================================================================

%----------------------------------------------------------------------------------------
\subsection{Integer Programming Mapping to QUBO}
\label{sec:methods:ilp}
%----------------------------------------------------------------------------------------

This section presents the formalism which maps a general ILP to a QUBO.

%----------------------------------------------------------------------------------------
\subsubsection{Integer linear programming}
\label{sec:methods:ilp-def}
%----------------------------------------------------------------------------------------
Formally, ILP is defined as the optimization of a linear function
\begin{align}
    \label{eq:initial-ip-def}
    \vec x_0 = \mathrm{arg} &\min\limits_{x}(\sum_i c_i x_i)
\end{align}
subject to
\begin{align}
    \label{eq:ilp-constraints}
    &\sum_i A_{ai}x_i +b_a \leq 0 \,,\\
    & x_i  \in \mathbbm{N}
\end{align}

where the natural numbers $\mathbbm{N}$ include zero, $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations.

%----------------------------------------------------------------------------------------
\subsubsection{Slack variables}
\label{sec:methods:slack}
%----------------------------------------------------------------------------------------

The inequality constraints in eq.~\eqref{eq:ilp-constraints} can be rewritten as an equality by introducing slack variables $s$ such that the ILP becomes
\begin{align}
    \vec x_0 = \mathrm{arg} &\min(\sum_i c_i x_i)\,,
\end{align}
subject to
\begin{align}& \sum_i A_{a i}x_i + s_a + b_a = 0,\\
    & s_a \in \mathbb{N},\\
    & x_i \in \mathbb{N}.
\end{align}
While the coefficients of the inequality constraints are not required to be integer valued, the inequalities can be trivially rescaled such that $s_i \in \mathbbm{N}$ given fixed precision coefficients $A_{ij}$ and $b_i$.
As a result, we can restrict both $x_i$ and $s_i$ to sample over only positive integer values as allowed by Eq.~(\ref{eq:int_to_bin}).

Because quantum annealing is capable of solving quadratic problems, we extend the proposed algorithm to solve integer quadratic optimization problems as well such that
\begin{align}
    \vec x_0 = \mathrm{arg}\min\limits_{x}\left(\sum_{ij} x_i d_{ij} x_j + \sum_i c_i x_i\right)
\end{align}
without the introduction of auxiliary qubits.

%----------------------------------------------------------------------------------------
\subsubsection{Bit basis}
\label{sec:methods:bit-basis}
%----------------------------------------------------------------------------------------

We map integer variables $\vec z = \vec x, \vec s$ to qubits under the following transformation~\cite{Chang:2018uoc}
\begin{align}
    z_i = & \sum_{r=0}^{R_i-1} 2^r \psi_{ri}
\label{eq:int_to_bin}
\end{align}
where $\psi_{ri} \in \{0, 1\}$.
The number of qubits used to represent the $i$-th integer variable is allowed to vary with $R_i$.
This transformation can be expressed as a rectangular matrix.
For example, a vector of two integer variables $z_0$ and $z_1$ represented by one and two qubits respectively is given as
\begin{align}
    \begin{pmatrix}
    z_0\\
    z_1
    \end{pmatrix}
    = &
    \begin{pmatrix}
    2^0 & 0 & 0\\
    0 & 2^0 & 2^1
    \end{pmatrix}
    \begin{pmatrix}
    \psi_{00}\\
    \psi_{01}\\
    \psi_{11}
    \end{pmatrix}
    \equiv T^z \begin{pmatrix}
    \psi_{00}\\
    \psi_{01}\\
    \psi_{11}
    \end{pmatrix}
\end{align}
If all variables are represented by the same number of qubits--$R_i$ is a constant for all $i$--then one can express the transformation as tensor product of bit vectors
\begin{align}
    \mathcal{R} = & \begin{pmatrix} 2^0 & \dots & 2^{R-1}\end{pmatrix},\\
    \mathcal{Z} = & \begin{pmatrix} z_0 & \dots & z_{N-1}\end{pmatrix},\\
    |\mathds{1}| = & |\mathcal{Z}|,\\
    T^z = & \mathds{1}\otimes \mathcal{R}.
\end{align}

%----------------------------------------------------------------------------------------
\subsubsection{Integer Programming Mapping QUBO}
\label{sec:methods:ilp-qubo}
%----------------------------------------------------------------------------------------

The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
    \chi^2 &= \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu \right.
    \nonumber\\ \label{eq:qubo-min}
    &\qquad\left.+ p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],\\
    \vec x_0 &=  T^x \vec \Psi^x_0 \, ,
    \\
    (\vec \Psi^x_0, \vec \Psi^s_0) &= \mathrm{arg}\min\limits_{\vec{\Psi}}(\chi^2) \,, &
\end{align}
where $p$ is an external parameter representing the strength of the penalty when violating constraints.
The objective function can be represented as a QUBO Hamiltonian
\begin{align}
    E = &
    \begin{pmatrix}
    \Psi^x & \Psi^s
    \end{pmatrix}
    \begin{pmatrix}
    Q_{xx} & Q_{xs}\\
    Q_{sx} & Q_{ss}
    \end{pmatrix}
    \begin{pmatrix}
    \Psi^x\\ \Psi^s
    \end{pmatrix} + pb^2,\\
    \equiv & \Psi^T Q \Psi + C,
    \label{eq:matrix_form}
\end{align}
where
\begin{align}
    Q_{xx} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x \nonumber \\
    &+ \mathrm{Diag}_{\psi^x}(c) T^x,\\
    Q_{xs} = & Q_{sx}^T = p {T^{x}}^T A^T T^s,\\
    Q_{ss} = & p\left[ {T^{s}}^T T^s + \mathrm{Diag}_{\psi^s}\left( {T^{s}}^T b + b^T T^s\right) \right].
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

%----------------------------------------------------------------------------------------
\subsubsection{Comment on penalty term}
\label{sec:methods:ilp-qubo-comments}
%----------------------------------------------------------------------------------------

The minimal energy solution to eq.~\eqref{eq:initial-ip-def} and eq.~\eqref{eq:qubo-min} are exactly the same if the penalty term is larger than the energy gap of the first excited solution: $p > E_1 - E_0$.
Thus some knowledge of the problem is required.
In principle, it is possible to set the penalty term arbitrarily large, at the cost of problem resolution: large $p$-values increase the highest available energy of the system by multiples of $p$.
After normalization of the QUBO, this corresponds to decreasing the energy gap between the ground state and the first excited state.
Thus, if solvers have finite precision, one must estimate reasonable values for $p$: for large $p$ more solutions fulfill the constraints, while for small $p$, more solutions minimize the objective function.

%----------------------------------------------------------------------------------------
\subsection{Example: Minimum Dominating Set Problem}
\label{sec:methods:mds}
%----------------------------------------------------------------------------------------

This section demonstrates the formalism in case of an example ILP: the Minimum Dominating Set problem.

%----------------------------------------------------------------------------------------
\subsubsection{Definition of the Minimum Dominating Set Problem}
\label{sec:methods:mds-def}
%----------------------------------------------------------------------------------------

Given a graph $G(E,V)$, defined by the set of $V$ vertices and $E$ edges, and a subset of vertices $D \subseteq V$.
If the set of all vertices not in $D$ is adjacent to at least one vertex in $D$, then $D$ is a dominating set.
This is equivalent to requiring the set of nearest-neighbor vertices of $D$ (exclusive) and $D$ cover all vertices $N(D) \cup D = V$ (an example is given by Fig.~\ref{fig:dominating_sets}a).
The set $D$ is a minimal dominating set if there is no proper subset of $D$ that is a dominating set, {\it{i.e.}}, the removal of any vertex in $D$ results in $N(D) \cup D  \neq V$.
An example is given by Fig.~\ref{fig:dominating_sets}b.
The domination number of $D$ is given by the cardinality of $|D| \equiv \overline{\overline{D}}$.
The minimum dominating set (MDS) is defined by $D$ with the smallest domination number.
Fig.~\ref{fig:dominating_sets}c shows an example of the minimum dominating set of $G(V, E)$ and is different from the minimal dominating set.
We emphasize that while the maximum independent set is always a minimal dominating set as exemplified by Fig.~\ref{fig:dominating_sets}b, the minimum dominating set, in general, can have a smaller domination number.
As a result, the solution to the dominating set problem can not be obtained by solving for the maximum independent set, a problem that is well studied for quantum annealers~\cite{}.

@@

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time.
For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.
We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.
\begin{figure*}
    \centering
    \begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
    \includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
    \includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
    \includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
    \includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
    \centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
    \end{tabular}
    \caption{
        Example of greedy MDS algorithm:
        The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
        Connected nodes are neighbors of the DS (turquoise) or already in the DS.
        In (a), one finds three maximally unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
        In the first step $a\to b$, randomly the node (4) is picked.
        In (b), one has three unconnected choices [(1), (2)] with each adjacent to another unconnected node and one node [(6)] which is only adjacent to neighbors of the DS.
        After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
        The found solution is a (minimal) dominating set but not a MDS.
        To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
    }
    \label{fig:mds-greedy}
\end{figure*}

%----------------------------------------------------------------------------------------
\subsubsection{Minimum Dominating Set Qubo}
\label{sec:methods:mds-qubo}
%----------------------------------------------------------------------------------------

The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
    f(x) = &\min\left(\sum_{\nu \in V} \psi^x_{\nu}\right),\\
    \textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1,\\
    & \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$.
For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
    f(x) = &\min(\sum_{\nu\in V} \psi^x_{\nu}),\\
    \textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0,\\
    & \psi_{\nu} \in \{0, 1\},\\
    & 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|\\
    & s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables.
Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

The target QUBO in the notation of eq.~\eqref{eq:matrix_form} reads
{\small
\begin{align}
    Q_{xx} &= \mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] \,,\\
    Q_{sx} &= - p(\mathbbm{1}+J^T)T^s\,,\\
    Q_{ss} &= p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]\,,\\
    C &=  p \overline{\overline{V}}\,,
\end{align}}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

%----------------------------------------------------------------------------------------
\subsection{Quantum Annealing}
\label{sec:methods:annealing}
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
\subsubsection{Modified annealing schedule}
\label{sec:methods:annealing-schedule}
%----------------------------------------------------------------------------------------

A quantum annealer evolves the ground state of an initial Hamiltonian into the ground state of the problem Hamiltonian by adiabatically changing the hardware which implements Hamiltonian.
Thus, the implemented Hamiltonian is explicitly time dependent
\begin{equation}
    H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
and $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, while $H^\textrm{problem}$ is given by Eq.~(\ref{eq:matrix_form}) for the MDS application.
The variable $s\in [0, 1]$ is the normalized annealing time and, in physical units, is allowed to vary between 1$\mu s$ to 2$ms$ on DWave.

On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits ($B(s) \to B_i(s)$).
In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in addition to the effects of applying positive and negative offsets.
Note that due to the way the offsets are implemented on the hardware level, systematic errors are introduced since the initial and final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets ($B_i(s) \neq B_j(s)$).
Fig.~\ref{fig:anneal_schedule} shows the maximum offset that we use for the study, which translates to approximately (due to non-linearity) a 5\% advancement (blue) or delay (red) in the schedule.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{./figures/anneal_schedule.pdf}
    \caption{asdf}
    \label{fig:anneal_schedule}
\end{figure}

>>>>

\documentclass[prd,twocolumn,tightenlines,preprintnumbers,showpacs,superscriptaddress,notitlepage,nofootinbib,eqsecnum,floatfix,longbibliography]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{bm,bbm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{dsfont}

@@

@@

\begin{abstract}
Abstract here...
\end{abstract}

@@

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tableofcontents

@@

\section{INTRODUCTION}
\label{sec:introduction}

Integer linear programming (ILP) is an integer optimization problem subject to inequality constraints and is a commonly tackled problem applicable to situations such as x~\cite{}, y~\cite{}, and z~\cite{}. In general however, ILP is classically NP-complete, and as a result, heuristic methods are employed~\cite{}. The NP-hardness of ILP can be understood by realizing that while the solution to an $n$-dimensional linear problem can be obtained in polynomial time, the optimal integer solution in general must be found in the $2^n$ integer solutions which surround the real number solution. Due to the importance and difficulty, in this work we present a method to obtain the optimal solution to ILP by employing methods of quantum annealing.

Quantum annealing solves a general quadratic binary optimization problem (QUBO) by slowly varying a time-dependent Hamiltonian~\cite{}. Through the adiabatic theorem of quantum mechanics, the annealer is prepared initially in a trivial ground state while the solution to ILP is encoded in the target Hamiltonian. Due to the explosion in research efforts towards hardware implementations of quantum annealers and future improvements to the annealing schedule~\cite{}, mapping ILP to QUBO provides a path forward towards obtaining optimal solutions to the class of integer optimization problems~\cite{2018Glover}.

Formally, ILP is defined as the optimization of a linear function
{\color{blue}
\begin{align}
f(x) = &\min\limits_{x}(\sum_i c_i x_i),\\
\textrm{subject to} \quad & \sum_i A_{ai}x_i +b_a \leq 0 \\
& x_i \geq 0,\\
& x_i \in \mathbb{Z}
\end{align}
}
where $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations. The inequality constraints can be rewritten as an equality by introducing slack variables $s$ such that the ILP becomes
{\color{blue}
\begin{align}
f(x) = &\min(\sum_i c_i x_i),\\
\textrm{subject to} \quad & \sum_i A_{a i}x_i + s_a + b_a = 0,\\
& s_a \geq 0,\\
& s_a \in \mathbb{Z},\\
& x_i \geq 0,\\
& x_i \in \mathbb{Z}.
\end{align}
}
Since quantum annealing solves the QUBO problem, we extend the proposed algorithm to solve integer quadratic optimization problems such that
{\color{blue}
\begin{align}
f(x) = \min\limits_{x}(\sum_{ij} x_i d_{ij} x_j + \sum_i c_i x_i)
\end{align}
}
without the introduction of auxiliary qubits.

{\color{red} Paragraph on how this is solved classically. Include strengths and drawbacks of classical solution.}

\section{RESULTS}
\label{sec:results}
We map integer variables $z_i$ to qubits under the following transformation~\cite{Chang:2018uoc}
\begin{align}
z_i = & \sum_{r=0}^{R_i-1} 2^r \psi_{ri}
\label{eq:int_to_bin}
\end{align}
where $\psi_{ri} \in \{0, 1\}$ while the number of qubits used to represent the $i$-th integer variable is allowed to vary with $R_i$. The transformation can in general be express as a rectangular matrix. For example, a vector of two integer variables $z_0$ and $z_1$ represented by one and two qubits respectively is given as
\begin{align}
\begin{pmatrix}
z_0\\
z_1
\end{pmatrix}
= &
\begin{pmatrix}
2^0 & 0 & 0\\
0 & 2^0 & 2^1
\end{pmatrix}
\begin{pmatrix}
\psi_{00}\\
\psi_{01}\\
\psi_{11}
\end{pmatrix}
\equiv T^z \begin{pmatrix}
\psi_{00}\\
\psi_{01}\\
\psi_{11}
\end{pmatrix}
\end{align}
The transformation can be reduced to a tensor product if $R_i$ is a constant for all $i$ such that
\begin{align}
\mathcal{R} = & \begin{pmatrix} 2^0 & \dots & 2^{R-1}\end{pmatrix},\\
\mathcal{Z} = & \begin{pmatrix} z_0 & \dots & z_{N-1}\end{pmatrix},\\
|\mathds{1}| = & |\mathcal{Z}|,\\
T^z = & \mathds{1}\otimes \mathcal{R}.
\end{align}

While the coefficients of the inequality constraints are not required to be integer valued, the inequalities can be trivially rescaled such that $s_i \in \mathbb{Z}$ given fixed precision coefficients $A_{ij}$ and $b_i$. As a result, we can restrict both $x_i$ and $s_i$ to sample over only positive integer values as allowed by Eq.~(\ref{eq:int_to_bin}).

\subsection{Integer Programming Mapping to QUBO}
\label{sec:results:ilp}
The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
\chi^2 = & \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu \right. \nonumber\\
&\left.+ p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],\\
\Psi = & \begin{pmatrix}
\Psi^x\\
\Psi^s
\end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization. The objective function can be represented as a QUBO Hamiltonian
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
Q_{xx} & Q_{xs}\\
Q_{sx} & Q_{ss} 
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + pb^2,\\
\equiv & \Psi^T Q \Psi + C,
\label{eq:matrix_form}
\end{align}
where
\begin{align}
Q_{xx} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x \nonumber \\
&+ \mathrm{Diag}_{\psi^x}(c) T^x,\\
Q_{xs} = & Q_{sx}^T = p {T^{x}}^T A^T T^s,\\
Q_{ss} = & p\left[ {T^{s}}^T T^s + \mathrm{Diag}_{\psi^s}\left( {T^{s}}^T b + b^T T^s\right) \right].
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

\subsection{Application to the Dominating Set Problem}
\label{sec:results:mds}

Given a graph $G(E,V)$, defined by the set of $V$ vertices and $E$ edges, and a subset of vertices $D \subseteq V$. If the set of all vertices not in $D$ is adjacent to at least one vertex in $D$, then $D$ is a dominating set. This is equivalent to requiring the set of nearest-neighbor vertices of $D$ (exclusive) and $D$ cover all vertices $N(D) \cup D = V$ (an example is given by Fig.~\ref{fig:dominating_sets}a). The set $D$ is a minimal dominating set if there is no proper subset of $D$ that is a dominating set, {\it{i.e.}}, the removal of any vertex in $D$ results in $N(D) \cup D  \neq V$. An example is given by Fig.~\ref{fig:dominating_sets}b. The domination number of $D$ is given by the cardinality of $|D| \equiv \overline{\overline{D}}$. The minimum dominating set (MDS) is defined by $D$ with the smallest domination number. Fig.~\ref{fig:dominating_sets}c shows an example of the minimum dominating set of $G(V, E)$ and is different from the minimal dominating set. We emphasize that while the maximum independent set is always a minimal dominating set as exemplified by Fig.~\ref{fig:dominating_sets}b, the minimum dominating set in general can have a smaller domination number. As a result, the solution to the dominating set problem can not be obtained by solving for the maximum independent set, a problem that is well studied for quantum annealers~\cite{}.

@@

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.
We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.
\begin{figure*}
	\centering
	\begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
	\end{tabular}
	\caption{
		Example of greedy MDS algorithm:
		The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
		Connected nodes are neighbors of the DS (turquoise) or already in the DS.
		In (a), one finds three unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
		In the first step $a\to b$, randomly the node (4) is picked.
		In (b), one has two unconnected choices [(1), (2)] with each adjacent to another unconnected node.
		After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
		The found solution is a (minimal) dominating set but not a MDS.
		To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
	}
	\label{fig:mds-greedy}
\end{figure*}

\subsubsection{Minimum Dominating Set}
The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
f(x) = &\min(\sum_{\nu \in V} \psi^x_{\nu}),\\
\textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1,\\
& \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$. For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
f(x) = &\min(\sum_{\nu\in V} \psi^x_{\nu}),\\
\textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0,\\
& \psi_{\nu} \in \{0, 1\},\\
& 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|\\
& s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables. Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

The target Hamiltonian can be written in the matrix form
\begin{widetext}
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
\mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] & - p(\mathbbm{1}+J^T)T^s\\
- p{T^s}^T(\mathbbm{1}+J)& p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + p \overline{\overline{V}}
\label{eq:matrix_form}
\end{align}
\end{widetext}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

\section{DISCUSSION}

\subsection{Finding the Dominating Set with a Quantum Annealer}
\label{sec:discussion:qa}
\begin{figure}[b]
	\centering
		\includegraphics[width=0.5\columnwidth]{./figures/linear_crop.pdf}
	\caption{Linear graphs $G(n)$ used in this study. Nodes denote vertices of the graphs and lines are undirected edges.}
	\label{fig:linear}
\end{figure}

For the following studies, we perform experiments on the \texttt{DW\_2000Q\_5} solver. The annealing time is optimized to 600$\mu$s after performing a study on the $G(6)$ graph as shown in Fig.~\ref{fig:at_scale}. In Fig.~\ref{fig:baseline}, we show results from the baseline experiment, without modification to the default DWave annealing schedule, and observe improvement over random guessing. However, the experiment results reflect the same degeneracy structure as random guessing, and scale exponentially poorly for larger number of vertices $n$. We explore one avenue towards improving the experiment results by introducing per-qubit annealing offsets into the time evolution.

\subsubsection{Modified annealing schedule}

In quantum annealing, the ground state of the problem Hamiltonian is obtained from the evolution of a time-dependent Hamiltonian,
\begin{equation}
H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
where $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, and $H^\textrm{problem}$ is given by Eq.~(\ref{eq:matrix_form}) for the MDS application. The variable $s\in [0, 1]$ is the normalized annealing time and is allowed to vary between 1$\mu s$ to 2$ms$ on the DWave. On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits. In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in addition to the effects of applying positive and negative offsets. Note that due to the way the offsets are implemented on the hardware level, systematic errors are introduced since the initial and final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets. Fig.~\ref{fig:anneal_schedule} shows the maximum offset that we use for the study, which translates to approximately (due to non-linearity) a 5\% advancement (blue) or delay (red) in the schedule.

\begin{figure}[b]
	\centering
	\includegraphics[width=\columnwidth]{./figures/anneal_schedule.pdf}
	\caption{asdf}
	\label{fig:anneal_schedule}
\end{figure}
